---
title: "angeliquelindberg_PeerCommentary_YYangEmily_01"
author: "Emily Yang"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1 - Working with R Data Structures

## Challenge 1 - Working with Vectors

### Creating Vector

```{r creating a vector}
v<-c("Stately","plump", "Buck", "Mulligan", "came", "from", "the", "stairhead", "bearing", "a", "bowl", "of", "lather", "on", "which", "a", "mirror", "and", "a", "razor", "lay", "crossed")
v
#could also have entered the whole sentence in quotes then used gsub("[[:punct:]]","",*string*)
```

### Extracting every third element

```{r splitting a vector}
v[seq(from = 1, to = 22, by = 3)]
#note that seq() may not always work well with character vectors
```

Side note: an important function is str()! For example,

```{r structure example}
str(v)
#so this is saying that v is a character vector with 22 values and then it lists them
```
Will provide another example of str() in challenge 5

## Challenge 2 - Working with matrices

### Creating a matrix

```{r creating matrix}
m<-matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m
```

### Subsetting a matrix

```{r subsetting by column}
X<-m[,c(2,3,6)]
#bracket notation allows for subsetting [row, column]; here I'm selecting multiple columns by combining them with the c funcion
X
```

```{r subsetting by row}
x<-m[c(6:8),]
x
```

```{r subsetting by row and column}
y<-m[2:6, 2:9]
y
```

## Challenge 3 - Working with arrays

### Creating an array

```{r creating an array}
a<-array(400:1, dim = c(5,5,4,4))
#note: this is a descending array, the dimensions are given by (row, column, # of matrices) 
a
```

### Subsetting an array

```{r subsetting array1}
a[1,1,1,2]
#this is selecting the element at the first row and first column of the 1,2 array
```

```{r subsetting array2}
a[2,3,2,]
#this is selecting the elements at the second row and third column of all arrays that start with 2
```

```{r subsetting array3}
a[1:5, 1:5, 3, 3]
#this is selecting the entire 3,3 matrix (because all five rows and all five columns)
```

## Challenge 4 - Working with lists

### Creating a list

```{r lemur and loris taxon}
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae","Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
```

```{r stepsirhine taxon}
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")
```

```{r tarsoid et al taxon}
Tarsoidea<-"Tarsiidae"
Catarrhini<- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Platyrrhini<- c("Cebidae", "Atelidae", "Pitheciidae")
```

```{r anthropoid taxon}
Anthropoidea<-list(Platyrrhini, Catarrhini)
names(Anthropoidea)<-list("Platyrrhini", "Catarrhini")
```

```{r haplorhine taxon}
Haplorhini<-list(Anthropoidea, Tarsoidea)
names(Haplorhini)<-list("Anthropoidea", "Tarsoidea")
```

```{r full primate}
Primates<-list(Haplorhini, Strepsirhini)
names(Primates)<-list("Haplorhini", "Strepsirhini")
Primates
```

## Challenge 5 - Working with data frames and conversion

### Making a matrix and then coercing it into a data frame
```{r creating matrix for df}
m3<-matrix(data = c(3,0,1,23,1,2,33,1,1,42,0,1,41,0,2), nrow = 5, ncol = 3, byrow = TRUE)
m3
```

```{r coercing into df}
df <- data.frame(m3)
df
```

### Coercing into different data classes
```{r coercing into character}
df[,2]<-as.character(df[,2])
```

```{r coercing into factor}
df[,3]<-as.factor(df[,3])
```
Side note 2: another str() function example:
```{r structure example2}
str(df)
```

